
---
title: "Recap of network exploration combinated with proto SSR"
author: "SC"
date: "`r Sys.time()`"
output: 
    bookdown::html_document2:
        code_folding: hide
        section_numbers: true
        toc: true
        toc_depth: 3
        toc_float: true
        fig_width: 10
        fig_height: 8

---


<style type="text/css">
body .main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
p.caption {
  font-size: 0.8em;
  font-weight: bold;
}

</style>

```{r}
knitr::opts_chunk$set(warning=FALSE,message=FALSE,cache=FALSE,collapse=TRUE,fig.show="hold",results="hide",fig.align="center")

library(sf)
library(igraph)
SSR=do.call("rbind",lapply(list.files("SSRgeoloc/",full.names = T),st_read))
plot(st_geometry(SSR))
sealip=st_read("Cartes/allseaborder_lowq.shp")
plot(st_geometry(st_crop(sealip,st_buffer(SSR,800000))),reset=F)
plot(st_geometry(SSR),add=T,pch=21,bg="red")

edgemat=matrix(0,nrow=nrow(SSR),ncol = nrow(SSR))
colnames(edgemat)=SSR$Name
rownames(edgemat)=SSR$Name
write.csv(file="edgematrix.csv",edgemat)
edgelist=cbind(expand.grid(SSR$Name,SSR$Name),0)
write.csv(file="edgelist.csv",edgelist)

netSSR=read.csv("data/SSR network - Matrix.csv",header=F)
rownames(netSSR)=netSSR[,1]
netSSR=netSSR[,2:ncol(netSSR)]
colnames(netSSR)=netSSR[1,]
netSSR=netSSR[2:nrow(netSSR),]
netSSR=as.matrix(netSSR)

#There is a rpoblem of names that have been swithced so:

dim(netSSR)
sum(!sapply(rownames(netSSR),function(i)i %in% colnames(netSSR)))
sum(!sapply(colnames(netSSR),function(i)i %in% rownames(netSSR)))
netSSR=netSSR[colnames(netSSR),]


netssr=igraph::graph_from_adjacency_matrix(netSSR)
V(netssr)$label.cex=.8
V(netssr)$label.color="black"
V(netssr)$size=.8
plot(netssr)

```


Detect missing things locations:

```{r}
missing=sapply(rownames(netSSR),function(i)i %in% SSR$Name)
missing[!missing]
missing=sapply(rownames(netSSR),function(i)i %in% SSR$Name)
coords=t(sapply(rownames(netSSR),function(i)st_coordinates(SSR[i == SSR$Name,])[1:2]))

SSR$geometry
plot(netssr)
```


We need to remobe the non coordinates things:


```{r}


par(mar=c(0,0,0,0))
netSSR=netSSR[!is.na(coords[,1]),!is.na(coords[,1])]
netssr=igraph::graph_from_adjacency_matrix(netSSR,mode = "undirected")
plot(st_geometry(st_crop(sealip,st_buffer(SSR,200000))),reset=F)
V(netssr)$size=25

V(netssr)$label=sapply(strsplit(V(netssr)$name,""),function(i)paste0(i[1:min(length(i),7)],collapse=""))
V(netssr)$label=gsub("\\W","",V(netssr)$label)
V(netssr)$label.dist=5
V(netssr)$color=adjustcolor("black",.4)
E(netssr)$color=adjustcolor("black",.4)
V(netssr)$color=adjustcolor("black",.4)
V(netssr)$label.color="black"
V(netssr)$label.cex=.85
V(netssr)$label.las=2
plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)
plot(netssr,layout=coords[V(netssr)$name,])

```


Then create network using the metal:

```{r}

source("tools.R")
KWPV=isotopes[ isotopes$Site.abreviation %in% c("NKH","NPW"),]
KWPV=KWPV[!(KWPV$LABEL.analytical %in%  c("SEALIP/TH/NPW/11","SEALIP/TH/NPW/1")),]
threshold  <- sapply(ratioiso,function(r){
                     v_ex=KWPV[,r]
                     alld=1-sapply(v_ex,function(i)i/v_ex)
                     quantile(alld[alld>0],probs=.75)[[1]]
})
ratioiso=c("X208Pb.204Pb","X207Pb.204Pb","X206Pb.204Pb")
isotopes=read.csv("data/DATA isotopie_Pryce_V13092022.csv")

isotopes$LABELS.shorts=sapply(lapply(strsplit(isotopes$LABEL.analytical,"/"),"[",3:4),paste0,collapse="/")

isotopes.consmat=createRatioMatrix(isotopes,ratioiso,threshold) 
dimnames(isotopes.consmat)=list(isotopes$LABELS.shorts,isotopes$LABELS.shorts)
isotopes.consmat[isotopes.consmat<3]=0
isotopes.consmat[isotopes.consmat==3]=1
isotopes.graph = graphFromAdj(isotopes.consmat,label=isotopes$LABELS.shorts)
isotopes.clus=cluster_leiden(isotopes.graph,objective_function="modularity")

isotopes.sitemat=createAreaAdjMat(isotopes,isotopes$Site.abreviation,isotopes.clus$membership,match=RGmatchNormalised)
isotopes.sitegraph = graphFromAdj(isotopes.sitemat,label=rownames(isotopes.sitemat))
alledges=E(isotopes.sitegraph)
tdl=alledges[ alledges$weight < .4]
isotopes.sitegraph=delete.edges(isotopes.sitegraph,tdl)
isotopes.sitecluster=cluster_leiden(isotopes.sitegraph,objective_function = "modularity")
V(isotopes.sitegraph)$color=isotopes.sitecluster$membership
E(isotopes.sitegraph)$width=E(isotopes.sitegraph)$weight
isotopes.sitegraph=colorEdges(isotopes.sitegraph)
plot(isotopes.sitegraph,main="networkd of sites")
isotopes.sitegraph.coords=coordForGraph(isotopes.sitegraph,isotopes,"Site.abreviation")
plot(isotopes.sitegraph.coords,ann=F,axes=F,type="n")
invalidcoords=apply(apply(isotopes.sitegraph.coords,2,is.na),1,any)

graph.sp=st_as_sf(as.data.frame(isotopes.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
plot(st_geometry(st_crop(sealip,graph.sp)),add=T)
plot(isotopes.sitegraph,layout=isotopes.sitegraph.coords,add=T,rescale=F)
plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)
```

Check community by communities

```{r,out.width="50%",fig.cap="communities detected for the whole dataset using normalized metric"}

for(gp in which(lengths(isotopes.sitecluster,use.names=F)>1)){
    sub=subgraph(isotopes.sitegraph,isotopes.sitecluster[[gp]])
    sub=set_graph_attr(sub,"layout",isotopes.sitegraph.coords[V(sub)$name,])
    try({
    plot(isotopes.sitegraph.coords,ann=F,axes=F,type="n")
    plot(sub,add=T,rescale=F)
    invalidcoords=apply(apply(isotopes.sitegraph.coords,2,is.na),1,any)
    graph.sp=st_as_sf(as.data.frame(isotopes.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
    plot(st_geometry(st_crop(sealip,graph.sp)),add=T)
    E(netssr)$width=5
    E(netssr)$color="red"
    plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)
    })
}

```

Using only  Bronze age artefact and removing links that are below a certain threshold

```{r,out.width="100%",fig.cap="Networks of sites with bronze age artefacts"}
tmpiso=isotopes[grep("Bronze",isotopes$Large.period),]

tmpiso.consmat=createRatioMatrix(tmpiso,ratioiso,threshold) 
dimnames(tmpiso.consmat)=list(tmpiso$LABELS.shorts,tmpiso$LABELS.shorts)
tmpiso.consmat[tmpiso.consmat<3]=0
tmpiso.consmat[tmpiso.consmat==3]=1
tmpiso.graph = graphFromAdj(tmpiso.consmat,label=tmpiso$LABELS.shorts)
tmpiso.clus=cluster_leiden(tmpiso.graph,objective_function="modularity")
tmpiso.sitemat=createAreaAdjMat(tmpiso,tmpiso$Site.abreviation,tmpiso.clus$membership,match=RGmatchNormalised)
tmpiso.sitegraph = graphFromAdj(tmpiso.sitemat,label=rownames(tmpiso.sitemat))

alledges=E(tmpiso.sitegraph)
tdl=alledges[ alledges$weight < .3]
tmpiso.sitegraph=delete.edges(tmpiso.sitegraph,tdl)

tmpiso.sitecluster=cluster_leiden(tmpiso.sitegraph,objective_function = "modularity")
V(tmpiso.sitegraph)$color="dark green"
V(tmpiso.sitegraph)$size=20
V(tmpiso.sitegraph)$label.dist=5
E(tmpiso.sitegraph)$width=E(tmpiso.sitegraph)$weight
tmpiso.sitegraph=colorEdges(tmpiso.sitegraph)
plot(tmpiso.sitegraph,main="networkd of sites")
tmpiso.sitegraph.coords=coordForGraph(tmpiso.sitegraph,tmpiso,"Site.abreviation")
plot(st_geometryst_buffer(netssr.sf,50000),ann=F,axes=F,type="n")
invalidcoords=apply(apply(tmpiso.sitegraph.coords,2,is.na),1,any)
graph.sp=st_as_sf(as.data.frame(tmpiso.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
netssr.sf=st_as_sf(as.data.frame(coords[!is.na(coords[,1]),]),coords=c(1,2))
plot(st_geometry(st_crop(sealip,st_buffer(netssr.sf,50000))),add=T)
plot(tmpiso.sitegraph,layout=tmpiso.sitegraph.coords,add=T,rescale=F)
plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)

```

Separate communities

```{r,out.width="50%",fig.cap="communities detected for the whole dataset using normalized metric"}

for(gp in which(lengths(tmpiso.sitecluster,use.names=F)>1)){
    sub=subgraph(tmpiso.sitegraph,tmpiso.sitecluster[[gp]])
    sub=set_graph_attr(sub,"layout",tmpiso.sitegraph.coords[V(sub)$name,])
    try({
        plot(tmpiso.sitegraph.coords,ann=F,axes=F,type="n")
        plot(sub,add=T,rescale=F)
        invalidcoords=apply(apply(tmpiso.sitegraph.coords,2,is.na),1,any)
        graph.sp=st_as_sf(as.data.frame(tmpiso.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
        plot(st_geometry(st_crop(sealip,graph.sp)),add=T)
        E(netssr)$width=5
        E(netssr)$color="red"
        plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)
    })
}


```

