---
title: "Isotpo"
author: "SC"
date: "`r Sys.time()`"
output: 
    bookdown::html_document2:
        code_folding: show
        section_numbers: true
        toc: true
        toc_depth: 3
        toc_float: true
---

# Isotope data

```{r,initialisation}
knitr::opts_chunk$set(warning=FALSE,message=FALSE,cache=FALSE,collapse=TRUE,fig.show="hold",results="hide")
```


```{r}
library(sf)
seashapefile=st_read("Cartes/allseaborder_lowq.shp")
seashapefile=st_geometry(seashapefile)

```

```{r}
pryce=read.csv("data/DATA isotopie_Pryce_V4.csv")
pryce[,c("Easting...Longitude","Northing.Latitude")]=apply(pryce[,c("Easting...Longitude","Northing.Latitude")],2,as.numeric)
spat=pryce[,c("Easting...Longitude","Northing.Latitude")]
spat=spat[!is.na(spat[,1]),]
p.sf=st_as_sf(spat,coords=c("Easting...Longitude","Northing.Latitude"),crs=st_crs(seashapefile))
seashapefile.cr=st_geometry(st_crop(seashapefile,st_buffer(p.sf,dist=100)))
pryce[,c("River.basin.coordinates.E","River.basin.coordinates.N")]=apply(pryce[,c("River.basin.coordinates.E","River.basin.coordinates.N")],2,as.numeric)
pryce=pryce[!is.na(pryce$River.basin.coordinates.E),]
ratioiso=c("X208Pb.204Pb","X207Pb.204Pb","X206Pb.204Pb")
pryce[,ratioiso]=apply(pryce[,ratioiso],2,as.numeric)
```
We get only some "clean" artefacts:


```{r}
cleanlist=pryce[pryce$Corrosion.products %in% c("Low","Ingot","Slag"),]
persite=table(cleanlist$Site.abreviation)
cleanlist=pryce[pryce$Site.abreviation %in% names(persite)[persite>9] & pryce$Corrosion.products %in% c("Low","Ingot","Slag"),]
sitecol=rainbow(length(unique(cleanlist$Site.abreviation)),alpha=.8)
names(sitecol)=unique(cleanlist$Site.abreviation)
plot(cleanlist[,ratioiso[3:2]],bg=sitecol[cleanlist$Site.abreviation],pch=20+as.numeric(as.factor(cleanlist$Country)))
plot(cleanlist[,ratioiso[c(3,1)]],bg=sitecol[cleanlist$Site.abreviation],pch=20+as.numeric(as.factor(cleanlist$Country)))
val=legend("topleft",fill=sitecol,legend=names(sitecol),title="Site",bty="n")
newxleg=val$rect$left+val$rect$w
newyleg=val$rect$top

legend(x=newxleg,y=newyleg,pch=20+1:length(unique(cleanlist$Country)),legend=levels(as.factor(cleanlist$Country)),title="Country",bty="n")
```

Using tsnea to separate artefacts

```{r}
library(Rtsne)
isotopes.tsne=Rtsne(as.matrix(cleanlist[,ratioiso]),pca=F)
plot(isotopes.tsne$Y[,1],isotopes.tsne$Y[,2],bg=sitecol[cleanlist$Site.abreviation],pch=20+as.numeric(as.factor(cleanlist$Country)))
#text(isotopes.tsne$Y[,1],isotopes.tsne$Y[,2],pryce[pryce$Corrosion.products %in% c("Low","Ingot","Slag"),"Site.abreviation"],cex=.6)
```


Faire les distance "à la oli" => groupe les artefact avec une difference <.1%


D'abord recalculer les difference maximum entre les ratio ingot/slag d'un même sites:

```{r,show="asis"}
samegroups=pryce[pryce$Artefact.type %in% c("Slag","Ingot"),]
#remove weirdos
samegroups=samegroups[samegroups$LABEL.analytical != "SEALIP/TH/NPW/1",]

minim=sapply(unique(samegroups$Site.abreviation),function(site){
    sitesratio=samegroups[samegroups$Site.abreviation == site,ratioiso]
    apply(sitesratio,2,ratioSpan)
}
)

(1-apply(minim,1,min))*100
knitr::kable(t(minim))

oliselection=minim[,c("PBL","NPW","NKH")]
knitr::kable(t(oliselection))
recomp=(1-apply(oliselection,1,min))*100
knitr::kable(recomp)
recomp=1-apply(oliselection,1,min)
```

Which correspond to oli's computation:
```{r,show="asis"}
olidiff=read.csv("data/prod sys var.csv",header=F)
knitr::kable(olidiff[6,c(4,6,8)]-recomp)
```

On va donc ensuite comparer chaque element deux a deux et voir quel ratios sont inferieur aux ratio min calculé precedement

Using this as a graph

```{r}
library(igraph)
adjmat=createRatioMatrix(cleanlist,ratioiso,recomp)

par(mar=c(0,0,0,0))
limited=adjmat
for(i in 0:3){
    graph=graphFromAdj(limited,mask=i,label=cleanlist$Site.abreviation)
    V(graph)$color=sitecol[cleanlist$Site.abreviation]
    plot(graph)
}
```

Let use this simplified matrix to detect groups
```{r,out.width="50%"}
par(mar=c(0,0,0,0))
allgraphs=lapply(0:3,function(m)graphFromAdj(limited,mask=m,label=cleanlist$Site.abreviation))
allcomus=lapply(allgraphs,cluster_louvain)
allgroups=sapply(allcomus,membership)
for(i in 1:4){
    graph=allgraphs[[i]]
    V(graph)$color=allgroups[,i]
    plot(graph)
}
knitr::kable(allgroups)
```


Play with full dataset:
```{r}
allsites=createRatioMatrix(pryce,ratioiso,recomp)
allgraphs=lapply(0:3,function(i)graphFromAdj(allsites,mask=i,label=pryce$Site.abreviation))
alcomu=sapply(allgraphs,cluster_louvain)
allgroups=sapply(alcomu,membership)
plot(pryce$Easting...Longitude,pryce$Northing.Latitude,col=alcomu[[1]]$membership,pch=20)
```

Create inter basin connections

```{r,out.width="50%"}
par(mar=c(0,0,0,0))
pryce=cbind(pryce,allgroups)
for(i in 1:4){
    matdim2=creatSiteAdjMat(pryce,pryce$River.basin,allgroups[,i],match=simpmatch)
    sites.graph=graph.adjacency(matdim2, mode = "undirected", weighted = TRUE, diag = FALSE)
    allcomu=cluster_louvain(sites.graph)
    V(sites.graph)$label=unique(pryce$River.basin)
    V(sites.graph)$color=allcomu$membership
    V(sites.graph)$label.cex=.7
    V(sites.graph)$size=8
    V(sites.graph)$label.color="black"
    E(sites.graph)$color=1
    E(sites.graph)$width=E(sites.graph)$weight/50
    #sites.graph=set_graph_attr(sites.graph,"layout",layout(c(pryce$River.basin.coordinates.E,pryce$River.basin.coordinates.N)))
    plot(sites.graph)
    #plot(seashapefile)
    #plot(sites.graph,layout=unique(cbind(pryce$River.basin.coordinates.E,pryce$River.basin.coordinates.N)),add=F)
}
```
Same with geoloc

```{r,out.width="50%"}
basinriv=pryce[!is.na(pryce$River.basin.coordinates.E) ,]
allsites=createRatioMatrix(basinriv,ratioiso,recomp)
allgraphs=lapply(0:3,function(i)graphFromAdj(allsites,mask=i,label=basinriv$Site.abreviation))
alcomu=sapply(allgraphs,cluster_louvain)
allgroups=sapply(alcomu,membership)
par(mar=c(0,0,0,0))
coordsites=t(sapply(V(sites.graph)$label,function(lbl)apply(unique(basinriv[basinriv$River.basin == lbl,c("River.basin.coordinates.E","River.basin.coordinates.N")]),2,mean)))
seashapefile.cr=st_crop(seashapefile,st_buffer(st_as_sf(as.data.frame(coordsites),coords=1:2,crs=st_crs(seashapefile)),dist=200000))
for(i in 1:4){
    matdim2=creatSiteAdjMat(basinriv,basinriv$River.basin,allgroups[,i],match=simpmatch)
    sites.graph=graph.adjacency(matdim2, mode = "undirected", weighted = TRUE, diag = FALSE)
    allcomu=membership(cluster_louvain(sites.graph))
    sites.graph=allgraphs[[i]]
    V(sites.graph)$label=unique(basinriv$River.basin)
    V(sites.graph)$color=allcomu
    V(sites.graph)$label.cex=.7
    V(sites.graph)$size=100
    V(sites.graph)$label.color="black"
    E(sites.graph)$color=allcomu[head_of(sites.graph,E(sites.graph))]

    E(sites.graph)$width=E(sites.graph)$weight/10
    plot(seashapefile.cr,reset=F)
    plot(sites.graph,layout=coordsites,add=T,rescale=F)
}


```



Using sites instead of River Basin

First we remove site with few artefacts:

```{r}
countsite=table(pryce$Site.abreviation)
bignames=names(countsite[countsite>2])
bigsite=pryce[!is.na(pryce$Easting...Longitude) ,]
bigsite=bigsite[bigsite$Site.abreviation %in% bignames ,]
```


```{r,out.width="50%"}
par(mar=c(0,0,0,0))
allmatdim=lapply(1:4,function(i)creatSiteAdjMat(bigsite,bigsite$Site.abreviation,allgroups[,i],match=simpmatch))
allgraphs=lapply(allmatdim,graph.adjacency, mode = "undirected", weighted = TRUE, diag = FALSE)
allclus=lapply(allgraphs,cluster_louvain)
allcommus=sapply(allclus,membership)
for(i in 1:4){
    sites.graph=allgraphs[[i]]
    V(sites.graph)$label=unique(bigsite$Site.abreviation)
    V(sites.graph)$color=allcommus[,i]
    V(sites.graph)$label.cex=.7
    V(sites.graph)$size=8
    V(sites.graph)$label.color="black"
    E(sites.graph)$color=1
    E(sites.graph)$width=E(sites.graph)$weight/50
    #sites.graph=set_graph_attr(sites.graph,"layout",layout(c(bigsite$River.basin.coordinates.E,bigsite$River.basin.coordinates.N)))
    plot(sites.graph)
    #plot(seashapefile)
    #plot(sites.graph,layout=unique(cbind(bigsite$River.basin.coordinates.E,bigsite$River.basin.coordinates.N)),add=F)
}
```

Same but with geoloc
```{r,out.width="50%"}
par(mar=c(0,0,0,0))
coordsites=t(sapply(V(sites.graph)$label,function(lbl)apply(unique(bigsite[bigsite$Site.abreviation == lbl,c("Easting...Longitude","Northing.Latitude")]),2,mean)))
seashapefile.cr=st_crop(seashapefile,st_buffer(st_as_sf(as.data.frame(coordsites),coords=1:2,crs=st_crs(seashapefile)),dist=200000))
for(i in 1:4){
    sites.graph=allgraphs[[i]]
    V(sites.graph)$label=unique(bigsite$Site.abreviation)
    V(sites.graph)$color=allcommus[,i]
    V(sites.graph)$label.cex=.7
    V(sites.graph)$size=100
    V(sites.graph)$label.color="black"
    E(sites.graph)$color=allcommus[head_of(sites.graph,E(sites.graph)),i]

    E(sites.graph)$width=E(sites.graph)$weight/80
    plot(seashapefile.cr,reset=F)
    plot(sites.graph,layout=coordsites,add=T,rescale=F)
}


```

# Data Elements

## Artefact Network


```{r}
datappm=read.csv("data/SEALIP_BROGLASEA data_compo_Pryce_V4.csv")

# set an arbitrary epsilon value and replace "tr"s and "0"s with it
epsilon=0.0001

listelements=c("Sb","Sn","Ag","Mo","Bi","Pb","Se","As","Au","Zn","Cu","Ni","Co")

datappm[,listelements]=apply(datappm[,listelements],2,as.numeric)
datappm[,listelements]=apply(datappm[,listelements],2,function(i){i[i=="tr"]=epsilon;as.numeric(i)})
#datappm[,listelements]=apply(datappm[,listelements],2,function(i){i[i=="nd"]=epsilon;as.numeric(i)})
datappm[,listelements]=apply(datappm[,listelements],2,function(i){i[is.na(i)]=epsilon;as.numeric(i)})
datappm[,listelements]=apply(datappm[,listelements],2,function(i){i[i==0]=epsilon;as.numeric(i)})

datappm$Easting...Longitude= as.numeric(gsub("°","",datappm$Easting...Longitude))
datappm$Northing.Latitude= as.numeric(gsub("°","",datappm$Northing.Latitude))

datappm=datappm[!is.na(datappm$Easting...Longitude),]

ppms=datappm[,listelements]

##detect dupliate and not artefact with nothing detected
artnull=which(apply(ppms,1,sd)==0)
ttt=dist(ppms)
ttt=as.matrix(ttt)
dupli=which(apply(ttt,2,function(i)any(sum(i==0)>1)))
datappm=datappm[-artnull,]
datappm=datappm[-dupli,]
ppms=ppms[-artnull,]
ppms=ppms[-dupli,]

ppms=as.data.frame(apply(ppms,2,log)) #log

```

### PCA - dimeension reduction

Compute PCA scores:

```{r pcafig,fig.cap="cluster using directly the PCA result",out.width="50%"}
# remove eelement poorly detected
#ppms=ppms[,apply(ppms,2,function(i)sum(i==log(epsilon)))<(.4*nrow(ppms))]

ppms.pca=prcomp(ppms)
ppms.dist=dist(ppms.pca$x) #compute distance between PCA scores



ppms.ew=1/(exp(as.matrix(ppms.dist)))
ppms.graph=graph.adjacency(ppms.ew, mode = "undirected", weighted = TRUE, diag = FALSE)
V(ppms.graph)$label=as.character(datappm$Country)
V(ppms.graph)$label.cex=.4
V(ppms.graph)$size=8
V(ppms.graph)$frame.color="black"
V(ppms.graph)$frame.width=.5
V(ppms.graph)$label.color=adjustcolor("black",.6)
E(ppms.graph)$color=NA
ppms.graph=set_graph_attr(ppms.graph,"layout",layout_with_fr(ppms.graph))
plot(ppms.graph)
```

Use this graph to  compute the communities

```{r louvain.new, results="asis",fig.cap="simple visualisation of the network with louvain communities. We label node given the site (left) and site + period (right)  where the artefact has been found.",out.width="50%"}
par(mar=c(0,0,0,0))
ppms.louvain=cluster_louvain(ppms.graph)
knitr::kable(sizes(ppms.louvain))
ppms.dist=dist(ppms.pca$x) #compute distance between PCA scores
ncommu=length(sizes(ppms.louvain))

plot(ppms.graph,vertex.color=rainbow(ncommu,alpha=.6)[ppms.louvain$membership],vertex.label=NA)

plot(ppms.graph,vertex.color=rainbow(ncommu,alpha=.6)[ppms.louvain$membership])

```



## Site Network

```{r}
joined=cbind(datappm,louvain=ppms.louvain$membership)
sites=unique(joined$Site)

matdim2=creatSiteAdjMat(joined,joined$Site,joined$louvain,match=simpmatch)

sites.graph=graph.adjacency(matdim2, mode = "undirected", weighted = TRUE, diag = FALSE)
V(sites.graph)$label=as.character(sites)
V(sites.graph)$label.cex=.7
V(sites.graph)$size=8
V(sites.graph)$label.color="black"
E(sites.graph)$color=NA
sites.graph=set_graph_attr(sites.graph,"layout",layout_nicely(sites.graph))

```

visualise  communities of sites

```{r,out.width="100%"}
par(mar=c(0,0,0,0))
sites.communities=cluster_louvain(sites.graph)
comus=sites.communities$memberships[1,]
E(sites.graph)$color=comus[head_of(sites.graph,E(sites.graph))]
V(sites.graph)$color=comus
plot(sites.graph)
coordsites=t(sapply(V(regions.graph)$label,function(lbl)apply(unique(joined[joined$Site == lbl,c("Easting...Longitude","Northing.Latitude")]),2,mean)))
seashapefile.cr=st_crop(seashapefile,st_buffer(st_as_sf(as.data.frame(coordsites),coords=1:2,crs=st_crs(seashapefile)),dist=200000))
V(sites.graph)$size=100
plot(seashapefile.cr,lwd=0.6,reset=F)
plot(sites.graph,layout=coordsites,rescale=F,add=T)


```

Try with communities of regions:

```{r}
joined=cbind(datappm,louvain=ppms.louvain$membership)
regions=unique(joined$Region)

matdim2=creatSiteAdjMat(joined,joined$Region,joined$louvain,match=simpmatch)

regions.graph=graph.adjacency(matdim2, mode = "undirected", weighted = TRUE, diag = FALSE)
V(regions.graph)$label=as.character(regions)
V(regions.graph)$label.cex=.7
V(regions.graph)$size=8
V(regions.graph)$label.color="black"
E(regions.graph)$width=E(regions.graph)$weight/4

```

Compute communities of regions:

```{r,out.width="100%"}
par(mar=c(0,0,0,0))
regions.communities=membership(cluster_louvain(regions.graph))
E(regions.graph)$color=regions.communities[head_of(regions.graph,E(regions.graph))]
E(regions.graph)$color=regions.communities[head_of(regions.graph,E(regions.graph))]
V(regions.graph)$color=regions.communities
plot(regions.graph)
V(regions.graph)$size=100
coordregions=t(sapply(V(regions.graph)$label,function(lbl)apply(unique(joined[joined$Region == lbl,c("Easting...Longitude","Northing.Latitude")]),2,mean)))
seashapefile.cr=st_crop(seashapefile,st_buffer(st_as_sf(as.data.frame(coordregions),coords=1:2,crs=st_crs(seashapefile)),dist=200000))
plot(seashapefile.cr,lwd=0.6,reset=F)
plot(regions.graph,layout=coordregions,rescale=F,add=T)
```

