
---
title: "Recap of network exploration combinated with proto SSR"
author: "SC"
date: "`r Sys.time()`"
output: 
    bookdown::html_document2:
        code_folding: hide
        section_numbers: true
        toc: true
        toc_depth: 3
        toc_float: true
        fig_width: 10
        fig_height: 8

---


<style type="text/css">
body .main-container {
max-width: 1800px;
margin-left: auto;
margin-right: auto;
}
p.caption {
  font-size: 0.8em;
  font-weight: bold;
}

</style>

# Community detection using the whole dataset

```{r}
knitr::opts_chunk$set(warning=FALSE,message=FALSE,cache=FALSE,collapse=TRUE,fig.show="hold",results="hide",fig.align="center")
```
```{r}

library(sf)
library(igraph)
SSR=do.call("rbind",lapply(list.files("SSRgeoloc/",full.names = T),st_read))
#plot(st_geometry(SSR))
sealip=st_read("maps/allseaborder_lowq.shp")
#plot(st_geometry(st_crop(sealip,st_buffer(SSR,800000))),reset=F)
#plot(st_geometry(SSR),add=T,pch=21,bg="red")

edgemat=matrix(0,nrow=nrow(SSR),ncol = nrow(SSR))
colnames(edgemat)=SSR$Name
rownames(edgemat)=SSR$Name
#write.csv(file="edgematrix.csv",edgemat)
edgelist=cbind(expand.grid(SSR$Name,SSR$Name),0)
#write.csv(file="edgelist.csv",edgelist)

netSSR=read.csv("data/SSR network - Matrix.csv",header=F)
rownames(netSSR)=netSSR[,1]
netSSR=netSSR[,2:ncol(netSSR)]
colnames(netSSR)=netSSR[1,]
netSSR=netSSR[2:nrow(netSSR),]
netSSR=as.matrix(netSSR)

#There is a rpoblem of names that have been switched so:

#dim(netSSR)
#sum(!sapply(rownames(netSSR),function(i)i %in% colnames(netSSR)))
#sum(!sapply(colnames(netSSR),function(i)i %in% rownames(netSSR)))
#netSSR=netSSR[colnames(netSSR),]


netssr=igraph::graph_from_adjacency_matrix(netSSR)
V(netssr)$label.cex=.8
V(netssr)$label.color="black"
V(netssr)$size=.8
#plot(netssr)

```


Detect missing things locations:

```{r,eval=F}
missing=sapply(rownames(netSSR),function(i)i %in% SSR$Name)
missing[!missing]
missing=sapply(rownames(netSSR),function(i)i %in% SSR$Name)

#plot(netssr)
```


We need to remobe the non coordinates things:


```{r}
coords=t(sapply(rownames(netSSR),function(i)st_coordinates(SSR[i == SSR$Name,])[1:2]))


par(mar=c(0,0,0,0))
netSSR=netSSR[!is.na(coords[,1]),!is.na(coords[,1])]
netssr=igraph::graph_from_adjacency_matrix(netSSR,mode = "undirected")
plot(st_geometry(st_crop(sealip,st_buffer(SSR,200000))),reset=F)
V(netssr)$size=25
#V(netssr)$size=betweenness(netssr)/20

V(netssr)$label=sapply(strsplit(V(netssr)$name,""),function(i)paste0(i[1:min(length(i),7)],collapse=""))
V(netssr)$label=gsub("\\W","",V(netssr)$label)
V(netssr)$label.dist=5
V(netssr)$color=adjustcolor("black",.4)
E(netssr)$color=adjustcolor("black",.4)
V(netssr)$color=adjustcolor("black",.4)
V(netssr)$label.color="black"
V(netssr)$label.cex=.85
V(netssr)$label.las=2
plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)

```


Then create network using metal isotpots similarities.:

```{r}

source("R/tools.R")
isotopes=read.csv("data/DATA isotopie_Pryce_V4.1.csv")
KWPV=isotopes[ isotopes$Site.abreviation %in% c("NKH","NPW"),]
KWPV=KWPV[!(KWPV$LABEL.analytical %in%  c("SEALIP/TH/NPW/11","SEALIP/TH/NPW/1")),]
IDVC=c( "SEALIP/LA/PBL/1", "SEALIP/LA/PBL/42", "SEALIP/LA/PBL/2", "SEALIP/LA/PBL/6", "SEALIP/LA/PBL/5", "SEALIP/LA/PBL/4", "SEALIP/LA/PBL/7", "SEALIP/LA/PBL/34", "SEALIP/LA/PBL/32", "SEALIP/LA/PBL/12", "SEALIP/LA/PBL/35", "SEALIP/LA/PBL/13", "SEALIP/LA/PBL/40", "SEALIP/LA/PBL/11", "SEALIP/LA/PBL/39", "SEALIP/LA/PBL/36", "SEALIP/LA/PBL/18", "SEALIP/LA/PBL/14", "SEALIP/LA/PBL/37", "SEALIP/LA/PBL/33", "SEALIP/LA/PBL/3", "SEALIP/LA/PBL/16", "SEALIP/LA/PBL/15", "SEALIP/LA/PBL/20", "SEALIP/LA/PBL/19", "SEALIP/LA/PBL/21", "SEALIP/LA/PBL/10", "SEALIP/LA/PBL/47")
VC=isotopes[ isotopes$LABEL.analytical %in% IDVC,]
ratioiso=c("X208Pb.204Pb","X207Pb.204Pb","X206Pb.204Pb")

thresholdMix  <- sapply(ratioiso,function(r){
                     kpwd=getAllDist(KWPV[,r])
                     VC=getAllDist(VC[,r])
                     allconsistent=c(VC,kpwd)
                     quantile(allconsistent,probs=.95)[[1]]
})

thresholdKPWDonly  <- sapply(ratioiso,function(r){
                     kpwd=getAllDist(KWPV[,r])
                     VC=getAllDist(VC[,r])
                     #allconsistent=kpwd
                     allconsistent=c(kpwd)
                     quantile(allconsistent,probs=.95)[[1]]
})
threshold  <- sapply(ratioiso,function(r){
                     kpwd=getAllDist(KWPV[,r])
                     VC=getAllDist(VC[,r])
                     allconsistent=c(kpwd,VC)
                     quantile(allconsistent,probs=.95)[[1]]
})



isotopes=isotopes[isotopes$Site.abreviation != "TLS",]
isotopes=isotopes[isotopes$Site.abreviation != "",]

isotopes$LABELS.shorts=sapply(lapply(strsplit(isotopes$LABEL.analytical,"/"),"[",3:4),paste0,collapse="/")

isotopes.consmat=createRatioMatrix(isotopes,ratioiso,threshold) 
dimnames(isotopes.consmat)=list(isotopes$LABELS.shorts,isotopes$LABELS.shorts)
isotopes.consmat[isotopes.consmat<3]=0
isotopes.consmat[isotopes.consmat==3]=1
isotopes.graph = graphFromAdj(isotopes.consmat,label=isotopes$LABELS.shorts)
isotopes.clus=cluster_leiden(isotopes.graph,objective_function="modularity")

isotopes.sitemat=createAreaAdjMat(isotopes,isotopes$Site.abreviation,isotopes.clus$membership,match=RGmatchNormalised)
isotopes.sitegraph = graphFromAdj(isotopes.sitemat,label=rownames(isotopes.sitemat))
#alledges=E(isotopes.sitegraph)
#plot(density(alledges$weight))
#tdl=alledges[ alledges$weight < .16]
#isotopes.sitegraph=delete.edges(isotopes.sitegraph,tdl)
isotopes.sitecluster=cluster_leiden(isotopes.sitegraph,objective_function = "modularity")
V(isotopes.sitegraph)$color=isotopes.sitecluster$membership
E(isotopes.sitegraph)$width=exp(E(isotopes.sitegraph)$weight)-1
isotopes.sitegraph=colorEdges(isotopes.sitegraph)
plot(isotopes.sitegraph,main="Network of sites based on shared artefacts from same family")
isotopes.sitegraph.coords=coordForGraph(isotopes.sitegraph,isotopes,"Site.abreviation")
plot(isotopes.sitegraph.coords,ann=F,axes=F,type="n")
invalidcoords=apply(apply(isotopes.sitegraph.coords,2,is.na),1,any)

graph.sp=st_as_sf(as.data.frame(isotopes.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
plot(st_geometry(st_crop(sealip,graph.sp)),add=T)
plot(isotopes.sitegraph,layout=isotopes.sitegraph.coords,add=T,rescale=F)
plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)
```

Check community by communities

```{r,out.width="50%",fig.cap="communities detected for the whole dataset using normalized metric"}

for(gp in which(lengths(isotopes.sitecluster,use.names=F)>1)){
    sub=subgraph(isotopes.sitegraph,isotopes.sitecluster[[gp]])
    sub=set_graph_attr(sub,"layout",isotopes.sitegraph.coords[V(sub)$name,])
    try({
    plot(isotopes.sitegraph.coords,ann=F,axes=F,type="n")
    plot(sub,add=T,rescale=F)
    invalidcoords=apply(apply(isotopes.sitegraph.coords,2,is.na),1,any)
    graph.sp=st_as_sf(as.data.frame(isotopes.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
    plot(st_geometry(st_crop(sealip,graph.sp)),add=T)
    E(netssr)$width=5
    E(netssr)$color="red"
    plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)
    })
}

```
# Bronze period only

We use only Bronze age artefacts;  and removing links that are below a certain threshold.

```{r,out.width="100%",fig.cap="Networks of sites with bronze age artefacts"}
tmpiso=isotopes[grep("Bronze",isotopes$Large.period),]

tmpiso.consmat=createRatioMatrix(tmpiso,ratioiso,threshold) 
dimnames(tmpiso.consmat)=list(tmpiso$LABELS.shorts,tmpiso$LABELS.shorts)
tmpiso.consmat[tmpiso.consmat<3]=0
tmpiso.consmat[tmpiso.consmat==3]=1
tmpiso.graph = graphFromAdj(tmpiso.consmat,label=tmpiso$LABELS.shorts)
tmpiso.clus=cluster_leiden(tmpiso.graph,objective_function="modularity")
tmpiso.sitemat=createAreaAdjMat(tmpiso,tmpiso$Site.abreviation,tmpiso.clus$membership,match=RGmatchNormalised)
tmpiso.sitegraph = graphFromAdj(tmpiso.sitemat,label=rownames(tmpiso.sitemat))

#alledges=E(tmpiso.sitegraph)
#tdl=alledges[ alledges$weight < .3]
#tmpiso.sitegraph=delete.edges(tmpiso.sitegraph,tdl)

tmpiso.sitecluster=cluster_leiden(tmpiso.sitegraph,objective_function = "modularity")
V(tmpiso.sitegraph)$color=tmpiso.sitecluster$membership
V(tmpiso.sitegraph)$size=20
#V(tmpiso.sitegraph)$size=betweenness(tmpiso.sitegraph)
E(tmpiso.sitegraph)$width=2*(exp(E(tmpiso.sitegraph)$weight)-1)
tmpiso.sitegraph=colorEdges(tmpiso.sitegraph)
plot(tmpiso.sitegraph,main="Network of sites based on shared artefacts from same family")
V(tmpiso.sitegraph)$label.dist=5
tmpiso.sitegraph.coords=coordForGraph(tmpiso.sitegraph,tmpiso,"Site.abreviation")
invalidcoords=apply(apply(tmpiso.sitegraph.coords,2,is.na),1,any)
fullgraph.sp=st_as_sf(as.data.frame(tmpiso.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
netssr.sf=st_as_sf(as.data.frame(coords[!is.na(coords[,1]),]),coords=c(1,2))
plot(st_buffer(fullgraph.sp,100000),lwd=0.0001)
plot(st_geometry(st_crop(sealip,st_buffer(netssr.sf,50000))),add=T,col=adjustcolor("pink",.2))
plot(tmpiso.sitegraph,layout=tmpiso.sitegraph.coords,add=T,rescale=F)
E(netssr)$width=1
plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)

```

We separate and show only network fo same communities

```{r,out.width="50%",fig.cap="communities detected for the whole dataset using normalized metric"}

for(gp in which(lengths(tmpiso.sitecluster,use.names=F)>1)){
    sub=subgraph(tmpiso.sitegraph,tmpiso.sitecluster[[gp]])
    sub=set_graph_attr(sub,"layout",tmpiso.sitegraph.coords[V(sub)$name,])
    try({
        plot(st_buffer(fullgraph.sp,100000),lwd=0.00001)
        E(sub)$width=3
        invalidcoords=apply(apply(tmpiso.sitegraph.coords,2,is.na),1,any)
        graph.sp=st_as_sf(as.data.frame(tmpiso.sitegraph.coords[!invalidcoords,]),coords=1:2,crs=st_crs(sealip))
        plot(st_geometry(sealip),col=adjustcolor("pink",.2),add=T)
        plot(sub,add=T,rescale=F)
        E(netssr)$width=1
        E(netssr)$color="red"
        plot(netssr,layout=coords[V(netssr)$name,],add=T,rescale=F)
        box(which="outer")
    })
}


```


Let's explore this Bronze age network link by link.
So let check how the 156 bronze age samples have been classify, and how this classification is distributed within sites:

```{r,results="asis"}
cat("## Classification of evry samples in each groups\n")
famcompo=tapply(tmpiso$LABEL.analytical,paste0("gpe",tmpiso.clus$membership),unique)
na=lapply(names(famcompo),function(g) cat(paste(g,":\n<br>\t",paste0(famcompo[[g]],collapse = " - "),"<br>\n")))
sitecompo=tapply(paste0("gpe",tmpiso.clus$membership),tmpiso$Site.abreviation,table)
cat("## Distribution of group for each sites\n")
for(g in names(sitecompo)){cat(paste("site",g,"is made of:\n<br>"));sapply(names(sitecompo[[g]]),function(n)cat(paste(sitecompo[[g]][n],"samples from ",n ,",")));cat("<br>\n============<br>\n ")}
```

We can thus compare that to the 

```{r,fig.cap="bilplot with colouring using family detected via the threshold computing",fig.width=40,fig.height=20}

par(mfrow=c(1,2))
gpe=paste0("gpe",tmpiso.clus$membership)
cnt=table(gpe)
ncol=length(cnt[cnt>1])
sitecol=adjustcolor(rainbow(ncol),.6)
names(sitecol)=names(cnt[cnt>1])
plot(tmpiso[,ratioiso[3:2]],bg=sitecol[gpe],pch=21,cex=2)
text(tmpiso[,ratioiso[c(3,2)]],tmpiso$LABELS.shorts,,cex=.6,pos=1)
legend("bottom",legend=names(sitecol),pt.bg=sitecol,pch=21,pt.cex=2,ncol=4)
plot(tmpiso[,ratioiso[c(3,1)]],bg=sitecol[gpe],pch=21,cex=2)
text(tmpiso[,ratioiso[c(3,1)]],tmpiso$LABELS.shorts,,cex=.6,pos=1)
legend("topleft",legend=names(sitecol),pt.bg=sitecol,pch=21,pt.cex=2,ncol=4)

```
