---
title: "isotope data"
author: "SC"
date: "`r Sys.time()`"
output: 
    bookdown::html_document2:
        code_folding: show
        section_numbers: true
        toc: true
        toc_depth: 3
        toc_float: true
---

# Isotope data

```{r}
knitr::opts_chunk$set(warning=FALSE,message=FALSE,cache=TRUE,collapse=TRUE,fig.show="hold",results="hide")

fun_range <- function(x,n=100) {                              # Create user-defined function
  (x - min(x)) / (max(x) - min(x)) * 100
}
simpmatch  <-  function(i,j){
    if(length(i)>length(j)){
        tmp=i
        i=j
        j=tmp
    }
    length(na.omit(match(i,j)))
}
source("R/tools.R")
library(sf)
```
Get all shapefiles
```{r,eval=F}
allshapefiles=lapply(country,function(m) tmpm=readOGR(dsn = "Cartes/", layer=paste0(m,"_adm0") ))
seashapefile=do.call("rbind",allshapefiles)
rmarkdown::render("myanmarIso.Rmd",output_dir="/var/www/html")
```

```{r}
seashapefile=st_read("Cartes/allsea.shp")
seashapefile=st_make_valid(seashapefile)
seashapefile=st_union(seashapefile)

```

```{r,out.width="50%"}
pryce=read.csv("data/DATA isotopie_Pryce_V4.csv")
myanmar=pryce[pryce$Country == "Myanmar",]
myanmar$Easting...Longitude= as.numeric(gsub("°","",myanmar$Easting...Longitude))
myanmar$Northing.Latitude= as.numeric(gsub("°","",myanmar$Northing.Latitude))
myanmar=myanmar[!is.na(myanmar[,6]),]#useless with the new cleaner dataset
ratioiso=c("X208Pb.204Pb","X207Pb.204Pb","X206Pb.204Pb")
myanmar[,ratioiso]=apply(myanmar[,ratioiso],2,as.numeric)
myanmar=myanmar[!is.na(myanmar[,15]),]#useless with the new cleaner dataset
isotopes=myanmar[,ratioiso]
myanmar.pca=prcomp(myanmar[,ratioiso])
myanmar.dist=dist(myanmar.pca$x)
myanmar.clust=hclust(myanmar.dist)
myanmar.kclust=cutree(myanmar.clust,k=10) #playing comparing community cluster vs simple hierarchical clustering
plot(myanmar.pca$x[,1:2],pch=21,bg=rainbow(10,alpha=.6)[myanmar.kclust])
text(myanmar.pca$x[,1],myanmar.pca$x[,2],myanmar$River.basin,pch=21,bg=rainbow(10,alpha=.6)[myanmar.kclust],cex=.4)


plot(myanmar.pca$x[,1:2],pch=21,bg=rainbow(10,alpha=.6)[myanmar.kclust])
text(myanmar.pca$x[,1],myanmar.pca$x[,2],myanmar$Site,pch=21,bg=rainbow(10,alpha=.6)[myanmar.kclust],cex=.4)
```

Crop shapefile:

```{r}

plot(st_geometry(seashapefile),reset=F)
points(myanmar$Easting...Longitude,myanmar$Northing.Latitude,pch=21,bg=rainbow(10,alpha=.6)[myanmar.kclust])

coords=colnames(myanmar)[6:7]

myan.sf=st_as_sf(myanmar,coords=coords,crs=st_crs(seashapefile))

bbox=st_bbox(myan.sf)
bigbb=st_bbox(st_buffer(st_as_sfc(bbox),600000))

bbox[["xmin"]]=bigbb[["xmin"]]
bbox[["xmax"]]=bigbb[["xmax"]]

windowmyan=st_crop(seashapefile,st_buffer(st_as_sfc(bbox),200000))
windowmyan=st_simplify(windowmyan,dTolerance=1000)


plot(st_geometry(windowmyan),reset=F)
points(myanmar$Easting...Longitude,myanmar$Northing.Latitude,pch=21,bg=rainbow(10,alpha=.6)[myanmar.kclust])

text(myanmar[,19],myanmar[,18],myanmar$Site,pch=21,bg=rainbow(10,alpha=.6)[myanmar.kclust],cex=.4)

```


Plot Isotopes ratio classic bi-plot:

```{r}

myanmar$kclust=myanmar.kclust 
qualities=myanmar

plot(qualities[,"X206Pb.204Pb"],qualities[,"X207Pb.204Pb"],pch=21,bg=rainbow(10,alpha=.6)[qualities$kclust],ylab="207/204",xlab="206/204")
#text(qualities[,"X206Pb.204Pb"],qualities[,"X207Pb.204Pb"],qualities$Site.abreviation,cex=.4)

plot(qualities[,"X208Pb.204Pb"],qualities[,"X207Pb.204Pb"],pch=21,bg=rainbow(10,alpha=.6)[qualities$kclust],xlab="208/204",ylab="207/204")
#text(qualities[,"X208Pb.204Pb"],qualities[,"X207Pb.204Pb"],qualities$Site.abreviation,cex=.4)

ql.pca=prcomp(qualities[,ratioiso])
ql.dist=dist(ql.pca$x)
ql.clust=hclust(ql.dist)
ql.kclust=cutree(ql.clust,k=5)

oligroup=read.csv("data/group differences.csv")
qualities$klcust=ql.kclust
plot(ql.pca$x[,1:2],pch=21,bg=rainbow(5)[ql.kclust])
text(ql.pca$x[,1],ql.pca$x[,2],qualities$Site.abreviation,cex=.6)
```


Tsnea and networks:
```{r,fig.height=10,fig.width=10}
nk=6

code=strsplit(myanmar[,1],"/")
code=sapply(code,function(s)paste0(s[3:4],collapse="/"))


library(Rtsne)
isotopes.tsne=Rtsne(isotopes,pca=F,perplexity=10)
isotopes.dist=dist(isotopes.tsne$Y)
isotopes.clust=hclust(isotopes.dist, method = "ward.D")
isotopes.kclust=cutree(isotopes.clust,k=nk) #playing comparing community cluster vs simple hierarchical clustering
plot(isotopes.tsne$Y[,1],isotopes.tsne$Y[,2],col=rainbow(nk)[isotopes.kclust],pch=20,xlab="PC1",ylab="PC2")
plot(isotopes.tsne$Y[,1],isotopes.tsne$Y[,2],col=rainbow(nk)[isotopes.kclust],pch=20,xlab="PC1",ylab="PC2",main="tSNE")

plot(isotopes.tsne$Y[,1],isotopes.tsne$Y[,2],col=rainbow(nk)[isotopes.kclust],pch=20,main="tSNE")
#text(isotopes.tsne$Y[,1],isotopes.tsne$Y[,2],myanmar$Site,cex=.5)
text(isotopes.tsne$Y[,1],isotopes.tsne$Y[,2],code,cex=.5)
```

Create network using tSNEA based distance

```{r}
library(igraph)
isotopes.ew=1/(exp(as.matrix(isotopes.dist)))
isotopes.graph=graph.adjacency(isotopes.ew, mode = "undirected", weighted = TRUE, diag = FALSE)
V(isotopes.graph)$label=as.character(myanmar$Site)
V(isotopes.graph)$label.cex=.4
V(isotopes.graph)$size=8
V(isotopes.graph)$frame.color="black"
V(isotopes.graph)$frame.width=.5
V(isotopes.graph)$label.color=adjustcolor("black",.6)
E(isotopes.graph)$color=NA
isotopes.graph=set_graph_attr(isotopes.graph,"layout",layout_with_fr(isotopes.graph))
plot(isotopes.graph)
```

```{r louvain,results="hold",fig.cap="simple visualisation of the network with louvain communities. We label node given the site (left) and site + period (right)  where the artefact has been found.",out.width="50%",fig.show="hold",eval=F}
par(mar=c(0,0,0,0))
isotopes.louvain=cluster_louvain(isotopes.graph)
#while(length(table(isotopes.louvain$memberships[1,]))!=10)isotopes.louvain=cluster_louvain(isotopes.graph)
#isotopes.louvain=multilevel.community(isotopes.graph)
counts=as.data.frame(sort(table(isotopes.louvain$membership)))
colnames(counts)=c("community","size")
knitr::kable(counts)
plot(isotopes.graph,vertex.color=rainbow(length(isotopes.louvain),alpha=.6)[isotopes.louvain$membership])

plot(isotopes.graph,vertex.color=rainbow(length(isotopes.louvain),alpha=.6)[isotopes.louvain$membership],vertex.label=paste(myanmar$Site,myanmar$relative.period,sep="-"))

plot(st_geometry(windowmyan),lwd=0.1)
points(myanmar$Easting...Longitude,myanmar$Northing.Latitude ,pch=21,bg=rainbow(length(isotopes.louvain),alpha=.6)[isotopes.louvain$membership])
```

## Site Network

```{r,eval=F}
joined=cbind(myanmar,louvain=isotopes.louvain$membership)
sites=unique(joined$Site)

system.time({
matdim1=sapply(sites,function(s1) sapply(sites,function(s2) simpmatch(joined$louvain[ joined$Site == s1],joined$louvain[ joined$Site == s2])))
})

system.time({
matdim2=matrix(NA,nrow=length(sites),ncol=length(sites))
for(s1 in 1:length(sites)){
    for(s2 in 1:length(sites)){
        c1=joined$louvain[ joined$Site == sites[s1]]
        c2=joined$louvain[ joined$Site == sites[s2]]
        matdim2[s1,s2]=simpmatch(c1,c2)
    }
}
})
sum(matdim2-matdim1)

sites.graph=graph.adjacency(matdim1, mode = "undirected", weighted = TRUE, diag = FALSE)
V(sites.graph)$label=as.character(sites)
V(sites.graph)$label.cex=.7
V(sites.graph)$size=8
V(sites.graph)$label.color="black"
E(sites.graph)$color=NA
sites.graph=set_graph_attr(sites.graph,"layout",layout_nicely(sites.graph))

```

visualise  communities of sites

```{r,out.width="100%"}
par(mar=c(0,0,0,0))
sites.communities=cluster_louvain(sites.graph)
comus=sites.communities$memberships[1,]

plot(sites.graph,vertex.color=rainbow(length(table(comus)),alpha=.6)[comus])
plot(st_geometry(windowmyan),lwd=0.1)
points(joined$Easting...Longitude,joined$Northing.Latitude ,pch=21,bg=rainbow(length(table(comus)),alpha=.6)[comus])

comus=sites.communities$memberships[nrow(sites.communities$memberships),]

plot(sites.graph,vertex.color=rainbow(length(table(comus)),alpha=.6)[comus])
plot(windowmyan,lwd=0.1)
points(joined$Easting...Longitude,joined$Northing.Latitude ,pch=21,bg=rainbow(length(table(comus)),alpha=.6)[comus])

```

```{r}

samegroups=pryce[pryce$Artefact.type %in% c("Slag","Ingot"),]
#remove weirdos
samegroups=samegroups[samegroups$LABEL.analytical != "SEALIP/TH/NPW/1",]

minim=sapply(unique(samegroups$Site.abreviation),function(site){
    sitesratio=samegroups[samegroups$Site.abreviation == site,ratioiso]
    apply(sitesratio,2,ratioSpan)
}
)

knitr::kable((1-apply(minim,1,min))*100 )
knitr::kable(t(minim))

oliselection=minim[,c("PBL","NPW","NKH")]
knitr::kable(t(oliselection))
recomp=(1-apply(oliselection,1,min))*100
knitr::kable(recomp)
recomp=1-apply(oliselection,1,min)
```

Which correspond to oli's computation:
```{r,results="asis"}
olidiff=read.csv("data/prod sys var.csv",header=F)
knitr::kable(olidiff[6,c(4,6,8)]-recomp)
```

On va donc ensuite comparer chaque element deux a deux et voir quel ratios sont inferieur aux ratio min calculé precedement

Using this as a graph

```{r}
sitecol=rainbow(length(unique(myanmar$Site.abreviation)),alpha=.8)
names(sitecol)=unique(myanmar$Site.abreviation)

library(igraph)
adjmat=createRatioMatrix(myanmar,ratioiso,recomp)

par(mar=c(0,0,0,0))
limited=adjmat
isograph=graphFromAdj(limited,mask=3,label=myanmar$Site.abreviation)
V(isograph)$color=sitecol[myanmar$Site.abreviation]
plot(isograph)
```

Let use this simplified matrix to detect groups based on the grouping criterion selection
```{r}
comuiso=cluster_louvain(isograph)
comuiso=comuiso$memberships[2,]
V(isograph)$color=comuiso
plot(isograph)
```

Using these group, compute graph between different level of grouping 

```{r}

groups=c("Site.abreviation","Region","River.basin")
rb=st_read("Cartes/riverbasin_sitelimited.shp")
rb=st_make_valid(rb)
rb$ID_RB=1:nrow(rb)
{
    s=myanmar
    par(mar=c(0,0,2,0))
    for(g in groups){
        nsite=table(s[[g]])
        #ss=s[s[[g]] %in% names(nsite[nsite>=3]),]
        ss=s
        ggraph=groupGraphIso(ss,coords,g,ratioiso,recomp)
        gcoor=graph_attr(ggraph,"layout")
        gcoor=st_as_sf(as.data.frame(gcoor),coords=1:2,crs=st_crs(windowmyan))
        plot(windowmyan,reset=F,main="Myanmar Network")
        if(g == "River.basin"){
            gcoor$ID_N=1:length(V(ggraph))
            inter=st_intersection(gcoor,rb[st_is_valid(rb),])
            plot(st_crop(rb[inter$ID_RB,],st_buffer(gcoor,dist=200000)),col=adjustcolor(categorical_pal(max(V(ggraph)$color))[V(ggraph)$color[inter$ID_N]],.7),add=T)
        }
        else{
            V(ggraph)$size=50
            V(ggraph)$label.size=.8
            V(ggraph)$cex=.8
        }
        plot(ggraph,add=T,rescale=F)
    }

    #par(mar=c(4,5,2,2))
    #sitecol=rainbow(length(unique(s$Site.abreviation)),alpha=.8)
    #names(sitecol)=unique(s$Site.abreviation)
    #plot(s[,ratioiso[3:2]],col=sitecol[s$Site.abreviation],pch=as.numeric(as.factor(s$Country)),main="Myanmar Network")
    #plot(s[,ratioiso[c(3,1)]],col=sitecol[s$Site.abreviation],pch=as.numeric(as.factor(s$Country)))
    #legend("topleft",pch=1:length(unique(s$Country)),legend=levels(as.factor(s$Country)),title="Country",bty="n")
}

```
plot the netowkr not on the map for easier visu 
```{r}

ggraph=groupGraphIso(myanmar,coords,"Site.abreviation",ratioiso,recomp)
V(ggraph)$size=10
V(ggraph)$label.cex=.8
plot(ggraph,layout=layout_with_fr(ggraph))
```
